CF-SuGAR Power Calculations

### 2025-03-26 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR power pretty much done (needs writeup though)

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
registry$bmi[registry$bmi >= 200] <- NA
# Write for SAS
write.csv(registry,
  file = "./CF-SuGAR/Background/Power Calculations/Data/registry.csv",
  row.names = F, na = ""
)
```

@@

### Plot

@@

@@

```{r}
mod1 <- lmer(bmi ~ Age_YrEnd + (1 | eDWID), data = registry)
summary(mod1,ddf="Kenward-Roger")
```

@@

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
registry$bmi[registry$bmi >= 200] <- NA
# A1c group
registry$a1c_group <- cut(registry$A_hgba1c, c(-Inf, 5.9, Inf))
registry$a1c_group <- relevel(registry$a1c_group, ref = "(5.9, Inf]")
# Remove people with CFRD
registry <- registry %>%
  filter(CFRD != "CFRD+") %>%
  arrange(eDWID, ReviewYear)
# Write for SAS
write.csv(registry,
  file = "./CF-SuGAR/Background/Power Calculations/Data/registry.csv",
  row.names = F, na = ""
)
# Proportion of good and bad A1c
# Overall
prop.table(table(registry$a1c_group))
# By person (take each person's worst)
t <- registry %>%
  group_by(eDWID) %>%
  summarise(
    a1c_group = ifelse(any(a1c_group == "(5.9, Inf]"), "(5.9, Inf]", "(-Inf,5.9]")
  )
prop.table(table(t$a1c_group))
```

@@

### Plots

#### Spaghetti

@@

#### BMI correlation matrix by visit number

```{r}
t <- registry %>%
  group_by(eDWID) %>%
  mutate(visit_number = row_number()) %>%
  filter(!is.na(bmi)) %>%
  select(eDWID, visit_number, bmi) %>%
  pivot_wider(names_from = visit_number, values_from = bmi)
head(t)
cor(data.matrix(t[, -c(1)]), use = "pairwise.complete.obs")
```

@@

```{r}
mod1 <- lmer(bmi ~ a1c_group + Age_YrEnd + (Age_YrEnd | eDWID),
  data = registry
)
summary(mod1, ddf = "Kenward-Roger")
```

@@

#### FEV1 correlation matrix by visit number

```{r}
t <- registry %>%
  group_by(eDWID) %>%
  mutate(visit_number = row_number()) %>%
  filter(!is.na(A_FEV1_pct_predicted)) %>%
  select(eDWID, visit_number, A_FEV1_pct_predicted) %>%
  pivot_wider(names_from = visit_number, values_from = A_FEV1_pct_predicted)
head(t)
cor(data.matrix(t[, -c(1)]), use = "pairwise.complete.obs")
```
